---
title: "Lab 02 - Plastic waste"
author: "Insert your name here"
date: "Insert date here"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

Remove this text, and add your answer for Exercise 1 here.

```{r plastic-waste-continent}
# insert code here
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.7)

```

### Exercise 2

```{r plastic-waste-density}
# insert code here
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.3)
```
We specify color and fill inside aes() because we want those attributes to vary according to the continent variable in your dataset and it tells ggplot2 to create a separate color/fill for each continent. However, we specify alpha outside aes() because we want the same transparency level for all density curves. Since it does not vary by a data variable, we treat it as a fixed characteristic of the geometry, instead of a mapped aesthetic.

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.

```{r plastic-waste-violin}
# insert code here

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = continent,
    y = plastic_waste_per_cap
  )
) +
  geom_violin()


```

Violin plots show the density of the data across different values, making it easier to spot multiple peaks or skewed distributions. The width at different points along the violin indicates how many data points fall in that region. Box plots, however, make it very clear what the quartile valeus are and highlights the presence of outliers. 


### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-mismanaged}
# insert code here

ggplot(
  data = plastic_waste,
  aes(
    x = plastic_waste_per_cap,
    y = mismanaged_plastic_waste_per_cap
  )
) +
  geom_point()

```

```{r plastic-waste-mismanaged-continent}
# insert code here

ggplot(
  data = plastic_waste,
  aes(
    x = plastic_waste_per_cap,
    y = mismanaged_plastic_waste_per_cap,
    color = continent
  )
) +
  geom_point()


```

```{r plastic-waste-population-total}
# insert code here

ggplot(plastic_waste, aes(x = total_pop, y = plastic_waste_per_cap)) +
  geom_point() +
  labs(
    title = "Plastic Waste per Capita vs. Total Population",
    x = "Total Population",
    y = "Plastic Waste per Capita"
  )


```

```{r plastic-waste-population-coastal}
# insert code here

ggplot(plastic_waste, aes(x = coastal_pop, y = plastic_waste_per_cap)) +
  geom_point() +
  labs(
    title = "Plastic Waste per Capita vs. Coastal Population",
    x = "Coastal Population",
    y = "Plastic Waste per Capita"
  )

```

Neither relationships seem to suggest strong linear associations. 

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r recreate-viz}
# insert code here

plastic_waste %>%
  mutate(coastal_pop_prop = coastal_pop / total_pop) %>%
  filter(plastic_waste_per_cap < 3) %>%
  ggplot(aes(
    x = coastal_pop_prop,
    y = plastic_waste_per_cap,
    color = continent
  )) +
  geom_point() +
  geom_smooth(color = "black") +
  scale_color_viridis_d() +
  labs(
    x = "Coastal population proportion (Coastal / total population)",
    y = "Plastic waste per capita ",
    color = "Continent",
    title = "Plastic waste vs. coastal population proportion",
    subtitle = "by continent"
  ) +
  theme_minimal()

```

Done!


